A GPU Sonar Simulator for Automatic Target Recognition

### 2014-04-29 UAM2014/Article/article.tex
la2pdf commit - 2014:04:29	14:02:09	(Tue29.Apr)

We have developed a simulator for synthetic aperture sonar (SAS) that creates object templates very quickly with the aid of the massive computing power in graphics processing units (GPUs). It uses OpenGL to render a optical 2D image and depth map of scene with a custom 3D model of the object and seafloor, and OpenCL to post-process this data to form sonar templates.

By coincidence our implementation is not so different from the SIGMAS+ simulator developed at the NATO Undersea Research Centre (NURC)~\cite{Coiras2009a, Coiras2009b}. They both take advantage of OpenGL to render a 2D image from 3D models, are designed for side-looking sonar, and assume a Lambertian scattering model. However, while SIGMAS+ is geared more towards realistically looking sonar images and includes effects such as noise, we do not. Also, SIGMAS+ sum images obtained from multiple OpenGL rendering passes to create the sonar image, while we render once and tailor the sonar image with the flexibility of OpenCL.

does not use OpenCL but a OpenGL a multi-pass technique, and includes ting  W
is not designed for creating object templates, and adds e.g. noise to the images for them to more closely

@@

Citations
\begin{itemize}
\item Midelfart/Midtgaard: Adaptive templates \cite{Midelfart2010}
\item Simulation of sidescan images (Bell thesis \cite{Bell1997}, pre-work by V.Blake \cite{Blake1993})
\end{itemize}

@@

OpenGL is a popular, well-matured, multi-platform application programming interface (API) for rendering 2D and 3D vector graphics.

@@

>>>>

We have developed a simulator for synthetic aperture sonar (SAS) that creates object templates very quickly with the aid of the massive computing power in graphics processing units (GPUs). It loads a 3D model of the sea floor and an object of interest into OpenGL, where emitted sound waves are modeled with a light source placed at the sonar transmit location. We assume rough, isotropic surfaces that reflects sound energy equally in all directions. This can be modeled with a Lambertian scattering model~\cite{Bell1997} where the intensity of the backscattered sound is said to depend only on the on incidence angle of the transmitted signal onto the model surface. When rendering OpenGL is set up to produce an optical 2D intensity image and depth map that reveals the distance from each pixels to the propagation axis. For maximum flexibility this data is finally combined with OpenCL to produce the image templates.

By coincidence our implementation is not so different from the SIGMAS+ simulator developed at the NATO Undersea Research Centre (NURC)~\cite{Coiras2009a, Coiras2009b}. They both take advantage of OpenGL to render a 2D image from 3D models, are designed for side-looking sonar, and assume a Lambertian scattering model. However, while SIGMAS+ is geared more towards realistically looking sonar images by including effects such as noise, we do not. Also, SIGMAS+ sum images obtained from multiple OpenGL rendering passes to create the sonar image, while we render once and tailor the sonar image with the flexibility of OpenCL.

@@

\begin{itemize}
\item Midelfart/Midtgaard: Adaptive templates \cite{Midelfart2010}
\item Simulation of sidescan images (Bell thesis \cite{Bell1997}, pre-work by V.Blake \cite{Blake1993})
\end{itemize}

@@

Creating a 2D view of an arbitrary complex 3D scene is a non-trivial matter. This is why we decided to use the core OpenGL pipeline to do this for us. OpenGL is a popular, well-matured, and multi-platform application programming interface (API) for rendering 2D and 3D vector graphics. It relieves us from the intricacies of handling vertices, faces and textures, as well as from projecting this onto a suitable 2D image plane. This section explains how we set up OpenGL to do this for us, and then proceeds to describe how we post process the OpenGL images with OpenCL to form the final sonar image.

@@

If we place the camera at the sonar and look in the direction of the image OpenGL will generate a image of all the pixels that are ``seen'' from the sonar. However, this can only provide us with optically looking images, while was not designed to does not produce camera where the sonar is and render the scene   handling computing the projection of a  and rather post-process its result with OpenCL to obtain
